---
title: "Richness & Diversity"
author: "Adam A Kemberling"
date: "July 11, 2018"
output: 
  flexdashboard::flex_dashboard:
    theme: yeti
    vertical_layout: fill
---


Page 1 - Simpson 1
=====================================  

    
Column {data-width=500}
-------------------------------------
    
### Model Summary - GLMM Simposon's Index ~ elevation + bogaert + (1 | crms site)
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Model Summary
```{r libraries and data, echo = F, message = F, warning = F}
library(readxl)
library(tidyverse)
library(vegan)
library(rgdal)
library(ggThemeAssist)
library(extrafont)
library(car)
library(lme4)
library(MuMIn)
library(afex)

full.t <- read_csv("all_transect_data.csv")

new.df <- filter(full.t, dist_from_edge_m != "full_transect", Simpson != 1)
new.df$crms_site <- factor(new.df$crms_site, levels = c('345', '311', '369'))
#change labels
new.df$crms_site <- str_c("0", new.df$crms_site)
new.df$crms_site <- factor(new.df$crms_site, levels = c('0345', '0311', '0369'))
#fragmentation as factor
new.df$bog.f <- parse_factor(as.character(round(as.numeric(as.character(new.df$bogaert)), digits = 2)), 
  levels = ,c(as.character(sort(unique(round(as.numeric(as.character(new.df$bogaert)), digits = 2)), decreasing = T))))

#scale continuous predictors for effect testing
new.df$bogaert_scaled <- (new.df$bogaert - mean(new.df$bogaert))/ sd(new.df$bogaert)
new.df$elevation_scaled <- (new.df$correct_elev_m - mean(new.df$correct_elev_m, na.rm = T))/ sd(new.df$correct_elev_m, na.rm = T)

mod1 <- lmer(Simpson ~  bogaert + (1 | crms_site), data = new.df)
summary(mod1)
```

AIC score
```{r, echo=F}
AIC(mod1)
```

R-squared values
```{r, echo=F}
r.squaredGLMM(mod1)

```



For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.




Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM

```{r, echo = F, warning = F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(Simpson ~ bogaert, data = new.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "Simpson Diversity Index Scores")
for (i in 1:length(unique(new.df$crms_site))) {
  abline(coef(mod1)$crms_site[i,1],coef(mod1)$crms_site[i,2], lty = i, lwd = 2)
}
```   
 
###  Likelihood Ratio test

The approach to a likelihood ratio test is to compare the full model to a reduced and null model using an anova.

Basically, the test compares the fit of two models. The null hypothesis is that the smaller model is the “best” model; It is rejected when the test statistic is large. In other words, if the null hypothesis is rejected, then the larger model is a significant improvement over the smaller one.

Keep in mind that LRT-based null hypothesis tests are conservative when the null value (such as σ2=0) is on the boundary of the feasible space; in the simplest case (single random effect variance), the p-value is approximately twice as large as it should be (Pinheiro and Bates 2000).

1. Test full model against model without random effect to test for its significance
```{r}
m1 <- lmer(Simpson ~  bogaert + (1 | crms_site), data = new.df)
m0 <- lm(Simpson ~ bogaert, new.df)
anova(m1,m0) ## two sequential tests
```

2. calculate p-values for fixed effects
```{r}
mixed(Simpson ~  bogaert_scaled + (1 | crms_site), data = new.df, method = 'LRT')
```








Page 2 - Richness 1
=====================================  

Column {data-width=500}
-------------------------------------
    
### Model Summary - GLMM Richness ~ bogaert + (1 | crms site)
    
```{r Richness mod1 setup, echo = F, message = F, warning = F}

mod2 <- lmer(Richness ~ bogaert + (1 | crms_site), data = new.df)
summary(mod2)
```

AIC score
```{r, echo=F}
AIC(mod2)
```

R-squared values
```{r, echo=F}
r.squaredGLMM(mod2)

```



For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.


Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM
```{r, Richness GLMM1, echo = F, warning = F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(Richness ~ bogaert, data = new.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "Species Richness")
for (i in 1:length(unique(new.df$crms_site))) {
  abline(coef(mod2)$crms_site[i,1],coef(mod2)$crms_site[i,2], lty = i, lwd = 2)
}
```   
 
###  Likelihood Ratio test

1. Random Effect
```{r, echo=FALSE}
m2 <- lmer(Richness ~ bogaert + (1 | crms_site) , data = new.df)
m0 <- lm(Richness ~ bogaert, new.df)
anova(m2,m0) ## two sequential tests
```


2. calculate p-values for fixed effects
```{r, echo=FALSE}
mixed(Richness ~  bogaert_scaled + (1 | crms_site), data = new.df, method = 'LRT')
```


For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.










Page 3 - Richness 3
=====================================  

Column {data-width=500}
-------------------------------------
    
### Model Summary - GLMM Richness ~ bogaert + (1 | crms site) famil = 'poisson'
    
```{r Richness mod2 setup, echo = F, message = F, warning = T}

mod3 <- glmer(Richness ~ bogaert + (1 | crms_site) , data = new.df, family = poisson)
summary(mod3)
```


AIC score
```{r, echo=F}
AIC(mod3)
```

R-squared values
```{r, echo=F}
r.squaredGLMM(mod3)

```

For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.

Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM
```{r, Richness GLMM2, echo = F, warning = F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(Richness ~ bogaert, data = new.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "Species Richness")
for (i in 1:length(unique(new.df$crms_site))) {
  abline(coef(mod3)$crms_site[i,1],coef(mod3)$crms_site[i,2], lty = i, lwd = 2)
}
```   
 
###  Likelihood Ratio test
1. Random effect
```{r, echo=FALSE}
m3 <- glmer(Richness ~ bogaert + (1 | crms_site) , data = new.df, family = poisson)
m0 <- lm(Richness ~ bogaert, new.df)
anova(m3,m0) 
```

2. calculate p-values for fixed effects
```{r, echo=FALSE}
mixed(Richness ~ bogaert_scaled + (1 | crms_site), data = new.df, method = 'LRT')
```


For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.



Page 4 - Elevation 1
=====================================  

Column {data-width=500}
-------------------------------------
    
### Model Summary - GLMM Marsh Elevation ~ bogaert + (1 | crms site)


```{r Elevation mod setup, echo = F, message = F, warning = F}

mod4 <- lmer(correct_elev_m ~ bogaert + (1 | crms_site) , data = new.df)
summary(mod4)
```

AIC score
```{r, echo=F}
AIC(mod4)
```

R-squared values
```{r, echo=F}
r.squaredGLMM(mod4)

```

For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.


Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM
```{r, Elevation GLMM, echo = F, warning = F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(correct_elev_m ~ bogaert, data = new.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "Marsh Elevation (m)")
for (i in 1:length(unique(new.df$crms_site))) {
  abline(coef(mod4)$crms_site[i,1],coef(mod4)$crms_site[i,2], lty = i, lwd = 2)
}
```   




###  Likelihood Ratio test

1. Random effect
```{r, echo=FALSE}
m4 <- lmer(correct_elev_m ~ bogaert + (1 | crms_site) , data = new.df)
m0 <- lm(correct_elev_m ~ bogaert, new.df)
anova(m4,m0) 
```

2. calculate p-values for fixed effects
```{r, echo=FALSE}
mixed(correct_elev_m ~ bogaert_scaled + (1 | crms_site), data = new.df, method = 'LRT')
```


For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.







Page 5 - Simpson 2
=====================================  

    
Column {data-width=500}
-------------------------------------
    
### Model Summary - GLMM Simposon's Index ~ elevation + bogaert + (1 | crms site)

```{r Simpson mod2 setup, echo = F, message = F, warning = F}

mod5 <- lmer(Simpson ~ correct_elev_m + bogaert + (1 | crms_site) , data = new.df)
summary(mod5)
```


AIC score
```{r, echo=F}
AIC(mod5)
```

R-squared values
```{r, echo=F}
r.squaredGLMM(mod5)

```

For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.

Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM
```{r, Simpson GLMM2, echo = F, warning = F, include=F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(correct_elev_m ~ bogaert, data = new.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "Simpson's Diversity Index")
for (i in 1:length(unique(new.df$crms_site))) {
  abline(coef(mod5)$crms_site[i,1],coef(mod5)$crms_site[i,2], lty = i, lwd = 2)
}
```   

###  Likelihood Ratio test

1. Random Effect
```{r, echo=FALSE}
m5 <- lmer(Simpson ~ correct_elev_m + bogaert + (1 | crms_site) , data = new.df)
m0 <- lm(Simpson ~ correct_elev_m + bogaert, new.df)
anova(m5,m0) 
```

2. calculate p-values for fixed effects
```{r, echo=FALSE}
mixed(Simpson ~ elevation_scaled + bogaert_scaled + (1 | crms_site) , data = new.df[is.na(new.df$correct_elev_m) != TRUE,], method = 'LRT')
```


For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.


Page 6 - Richness 3
=====================================  

    
Column {data-width=500}
-------------------------------------
    
### Model Summary - Richness ~ elevation + bogaert + (1 | crms site)

```{r Richness mod3 setup, echo = F, message = F, warning = F}

mod6 <- lmer(Richness ~ correct_elev_m + bogaert + (1 | crms_site) , data = new.df)
summary(mod6)
```


AIC score
```{r, echo=F}
AIC(mod6)
```

R-squared values
```{r, echo=F}
r.squaredGLMM(mod6)

```

For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.


Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM
```{r, Richness GLMM3, echo = F, warning = F, include=F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(correct_elev_m ~ bogaert, data = new.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "Species Richness")
for (i in 1:length(unique(new.df$crms_site))) {
  abline(coef(mod6)$crms_site[i,1],coef(mod6)$crms_site[i,2], lty = i, lwd = 2)
}
```   

###  Likelihood Ratio test

1. Random Effect
```{r, echo=FALSE}
m6 <- lmer(Richness ~ correct_elev_m + bogaert + (1 | crms_site) , data = new.df)
m0 <- lm(Richness ~ correct_elev_m + bogaert, new.df)
anova(m6,m0) 
```

2. calculate p-values for fixed effects
```{r, echo=FALSE}
mixed(Richness ~ elevation_scaled + bogaert_scaled + (1 | crms_site) , data = new.df[is.na(new.df$correct_elev_m) != TRUE,], method = 'LRT')
```



For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.








Page 7 - %SI
=====================================  

Column {data-width=500}
-------------------------------------
    
### Model Summary - GLMM %SI ~ bogaert + (1 | crms site)
```{r SI mod3 setup, echo = F, message = F, warning = F}
water.df <- read_xlsx("noaa_blu_objective1_openwater_survey.xlsx")

#make light attenuation measurement 
sp <- .15
water.df <- water.df %>% mutate(
  month = ifelse(lubridate::month(date) == 4, "April", "September"),
  sal = as.numeric(sal),
  correct_elev_m = as.numeric(correct_elev_m),
  crms_site = factor(crms_site, levels = c("369", "311", "345")),
  frag = toupper(frag),
  subsite_id = str_c(crms_site, frag, sep = '-'),
  depth_cm = as.numeric(depth_cm),
  top_par1 = as.numeric(top_par1),                                          #bulb 1 is top bulb
  top_par2 = as.numeric(top_par2),                                          #bulb 2 is low bulb
  bot_par1 = as.numeric(bot_par1),
  bot_par2 = as.numeric(bot_par2),                                                                       
  Kd_surface = -(log10(top_par1 / top_par2)) / sp,                          #attenuation per .15m for surface measurements
  Kd_bottom = -(log10(bot_par1 / bot_par2)) / sp,                           #Kd at bottom, per .15m                     #multiply light lost in 15cm by (total depth/15cm increments)
  SI_full = (bot_par2/top_par1) * 100) %>% 
  filter(month == "April") 

bog.df <- structure(list(subsite_id = c("369-L", "369-M", "369-H", "311-L", "311-M", "311-H", "345-L", "345-M", "345-H"), 
                         bogaert = c(157.409232427562,158.794835413732, 144.600763346679, 152.825951296738, 147.180047035677, 
                                     139.741186393264, 145.714196114844, 148.353319527527, 139.722070072008)), 
                    class = "data.frame", row.names = c(NA, -9L))

#added bogaert scores
water.df <- left_join(water.df, bog.df, by = "subsite_id")
water.df$SI <- water.df$SI_full/100

#Standardize covariates for effect size testing
water.df$bogaert_scaled <- (water.df$bogaert - mean(water.df$bogaert, na.rm = T))/ sd(water.df$bogaert, na.rm = T)

#rename crms sites
#change labels
water.df$crms_site <- str_c("0", water.df$crms_site)
water.df$crms_site <- factor(water.df$crms_site, levels = c('0345', '0311', '0369'))
```


```{r, echo = FALSE, warning=FALSE}
#make the model
mod7 <- lmer(SI_full ~ bogaert + (1 | crms_site) , data = water.df)
summary(mod7)
```
AIC Score
```{r, echo=FALSE}
AIC(mod7)
```

R-squared values
```{r, echo=FALSE}
r.squaredGLMM(mod7)


```

Column {data-width=500}
-------------------------------------
   
### Plotting the GLMM
```{r, echo = F, warning = F, include=F}
par(mar = c(5,5,4,2), ps = 16, cex.lab = 1.125, cex.axis = 1, family = "Arial")
plot(SI_full ~ bogaert, data = water.df, las = 1,
     xlab = "Bogaert Fragmentation Metric", ylab = "% Surface Irradiance at Bottom")
for (i in 1:length(unique(water.df$crms_site))) {
  abline(coef(mod7)$crms_site[i,1],coef(mod7)$crms_site[i,2], lty = i, lwd = 2)
}
```   

###  Likelihood Ratio test

1. Random Effect
```{r, echo=FALSE}
m7 <- lmer(SI_full ~ bogaert + (1 | crms_site) , data = water.df)
m0 <- lm(SI_full ~  bogaert, water.df)
anova(m7,m0) 
```

2. calculate p-values for fixed effects
```{r, echo=FALSE}
mixed(SI_full ~ bogaert_scaled + (1 | crms_site) , data = water.df, method = 'LRT')
```


For some reason that I don't understand these sections won't scroll down unless I add more text here and I don't want to waste time troubleshooting it so I'm just going top do add this everywhere at the bottom of the codeblocks.